---
title: "Data Cleaning"
---

&nbsp;

# Dependency

```{r, message=FALSE, warning=FALSE, class.source = 'fold-show'}
library(tidyverse)
```

&nbsp;

# Introduction

The dataset used in this analysis is the [County Demographic Information (CDI)]("./data/cdi.csv") dataset. This dataset collected several features related to geographical locations, percent of populations by age, education and income levels, number of serious crimes, etc, of 440 U.S. counties from 1990-1992:

- `id`:       1-440

- `cty`:      Text string containing name of county

- `state`:    Two-letter text string containing abbreviation of state

- `area`:     Land area measured in square miles

- `pop`:      Estimated 1990 population

- `pop18`:    Percent of total population in age range from 18-34

- `pop65`:    Percent of total population in aged 65 or older

- `docs`:     Number of professionally active nonfederal physicians,1990

- `beds`:     Total number of beds, cribs, and bassinets during 1990

- `hsgrad`:   Percent of persons 25 years old or older who completed 12 or more years of school

- `bagrad`:   Percent of persons 25 years old or older with bachelor’s degrees

- `poverty`:  Percent of 1990 total population with income below poverty level

- `unemp`:    Percent of labor force that is unemployed

- `pcincome`: Income (in dollars) per person among those in 1990 population

- `totalinc`: Total personal income (in millions of dollars) among those in 1990 total population

- `region`:   Classification (according to U.S. Census Bureau) of region of the U.S. (1=Northeast, 2=North Central, 3=South, 4=West)

- `crimes`:   Total number of serious crimes in 1990, including murder, rape, robbery, aggravated assault, burglary, larceny-theft, and motor vehicle theft, as
reported by law enforcement agencies

&nbsp;

# Data Quality Assessment

Data quality assessment was performed to identifying issues in the dataset.

```{r, message=FALSE}
path = "./data/cdi.csv"
cdi = read_csv(path)
```
&nbsp;

## Completeness

The dataset is complete:

```{r}
skimr::skim(cdi)
```

&nbsp;

## Uniqueness

- There are no duplicate rows in this dataset:

```{r, class.source = 'fold-show'}
nrow(cdi) == nrow(unique(cdi))
```

&nbsp;

- However, some counties may share the same name:

```{r}
cdi %>% 
  left_join(cdi, by = "cty") %>% 
  filter(id.x != id.y) %>% 
  dplyr::select(cty, id.x, state.x, id.y, state.y) %>% 
  head(10) %>% 
  kableExtra::kable() %>% 
  kableExtra::kable_styling()
```

- If we combine the county name with its state name, there are still a few duplicated name:

```{r}
cdi = 
  cdi %>% 
  mutate(
    cty = map2_chr(as.character(state), cty, ~str_glue(.x, .y, .sep = "_"))
  )

cdi %>% 
  left_join(cdi, by = "cty") %>% 
  filter(id.x != id.y) %>% 
  dplyr::select(cty, id.x, state.x, id.y, state.y) %>% 
  kableExtra::kable() %>% 
  kableExtra::kable_styling()
```

- We can manually rename them:

```{r}
cdi[28, "cty"] = "MO_St._Loui_1"
cdi[123, "cty"] = "MO_St._Loui_2"
cdi[50, "cty"] = "MD_Baltimor_1"
cdi[57, "cty"] = "MD_Baltimor_2"
```



# Data Cleaning

The data cleaning process cleans the data as follows:

- factorized categorical variables.

- created a new variable, crm_1000, as the outcome, which describes the crime rate per 1000 population.

- created two variables that show the number of active physicians per 1000 people (docs_1000) and the number of hospital beds per 1000 people (beds_1000), considering the number of active physicians and the number of hospital beds may link to the total population in each county

- divided pop and pcincome by 1000 to avoid large differences among scales of features and too small coefficient in following regression

- removed variables that are inappropriate to serve as predictors, including the ID and name for each county and the total number of serious crimes reported.


```{r}
cdi = cdi %>% 
  mutate(
    id = as.factor(id),
    state = as.factor(state),
    region = as.factor(region),
    region = fct_recode(region,
      `Northeast` = "1",
      `North central` = "2",
      `South` = "3",
      `West` = "4"
    )
  ) %>% 
  mutate(
    crm_1000 = (crimes / pop) * 1000,
    docs_1000 = (docs / pop) * 1000,
    beds_1000 = (beds / pop) * 1000
  ) %>% 
  mutate(
    pop = pop / 1000,
    pcincome = pcincome / 1000
    ) %>% 
  # for regression analysis
  dplyr::select(-id, -cty, -crimes)
```


