---
title: "Exploratory Data Analysis"
---

The goal of EDA is to explore the relationship among variables and remove features improper for constructing the regression model.

&nbsp;

# Dependencies

```{r, message=FALSE, warning=FALSE, class.source = 'fold-show'}
library(tidyverse)
library(corrplot)
```

```{r, include=FALSE}
path = "./data/cdi.csv"
cdi = read_csv(path)

cdi = 
  cdi %>% 
  mutate(
    cty = map2_chr(as.character(state), cty, ~str_glue(.x, .y, .sep = "_"))
  )

cdi = cdi %>% 
  mutate(
    id = as.factor(id),
    state = as.factor(state),
    region = as.factor(region),
    region = fct_recode(region,
      `Northeast` = "1",
      `North central` = "2",
      `South` = "3",
      `West` = "4"
    )
  ) %>% 
  mutate(
    crm_1000 = (crimes / pop) * 1000,
    docs_1000 = (docs / pop) * 1000,
    beds_1000 = (beds / pop) * 1000
  ) %>% 
  mutate(
    pop = pop / 1000,
    pcincome = pcincome / 1000
    ) %>% 
  # for regression analysis
  dplyr::select(-id, -cty, -crimes)
```

&nbsp;

# Distribution of All Variables

The distributions of some variables follow a skewed normal pattern, while the others follow a Poisson pattern.

```{r, fig.height=8}
cdi %>%
  keep(is.numeric) %>% 
  gather() %>% 
  ggplot(aes(value)) +
    facet_wrap(~ key, scales = "free", ncol = 3) +
    geom_histogram()
```

&nbsp;

# Pairwise Relationship between Variables

Through inspecting the pairwise relationship among variables, we observed that some variables (`pop`, `totalinc`, `docs`, `beds`) has very large correlation coefficient. Thus, we remove these variables (except `pop`) to avoid introducing collinearity in further regression. 

&nbsp;

## Pairwise Relationship

```{r, fig.height=6}
pairs(cdi %>% dplyr::select(pop18, docs_1000, beds_1000, crm_1000, poverty, pcincome, hsgrad, bagrad, pop65, unemp, area, pop, totalinc, docs, beds), 
          lower.panel = panel.smooth, 
          upper.panel = NULL,
      font.labels = 3,
      cex.labels = 1,
      gap = 0)
```

&nbsp;

## Pairwise Correlation Coefficient:

```{r, fig.height=6}
col = colorRampPalette(c("#BB4444", "#EE9988", "#FFFFFF", "#77AADD", "#4477AA"))


corrplot(cor(keep(cdi, is.numeric)), 
         type = "lower", 
         method = "shade", 
         shade.col = NA, 
         tl.col = "black", 
         tl.srt = 45,
         col = col(200), 
         addCoef.col = "black", 
         cl.pos = "n",  
         order = "hclust", 
         number.cex = .8,
         number.font = 1, 
         diag = FALSE, 
         tl.pos = "ld", 
         tl.cex = .8)
```


&nbsp;

# Crime Rate Among State

We further observed the trends between crime rate and geographical locations/regions, and 
identified `region` instead of `state` as a proper predictor. The figure below displayed
a clear trend between county region and crime rate, suggest that `region` is a proper predictor.
Since `region` and `state` are both variables depicting geographical information and `state` has
too many levels as a categorical variable, we remove state.

```{r}
cdi %>% 
  ggplot(aes(x = reorder(state, -crm_1000), y = crm_1000, color = region)) +
  geom_point(alpha = .8) + 
  theme_classic() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5)) +
  theme(
    panel.grid.major.y = element_blank(),
    panel.grid.minor.y = element_blank(),
    panel.grid.major.x = element_blank(),
    panel.grid.minor.x = element_blank()
    #axis.text = element_text(size=12),
    #axis.title = element_text(size=15),
    #legend.text = element_text(size=12),
    #legend.title = element_text(size=15)
  ) + 
  scale_y_sqrt() +
  labs(y = "crime rate", x = "state")
```





